```{r}
accidents<-read.csv("https://www.dropbox.com/s/n5afymm63p4uq47/2015_accidents.csv?dl=1",header=T, stringsAsFactors=T)
consumption<-read.csv("https://www.dropbox.com/s/kf3tnzki29n11p2/consumptionDM.csv?dl=1",header=T,stringsAsFactors = T)
```

```{r}
head(accidents)
```

```{r}
head(consumption)
```

```{r}
library(dplyr)
library(stringr)
```
```{r}
#lapply takes a list and a function name and applies the 
#function to each element in the list. It returns the result in a list the same length as the input list.
#For example to know the class (i.e.  data type) of each column in the accidents data frame

lapply(accidents, class)
```

```{r}
#same as above but simplfies the resul into a vertor 
sapply(accidents, class)
```

```{r}
rd_lengths <- sapply(accidents$Response.difference, str_length)


```



```{r}
head(rd_lengths, 15)
```
```{r}
typeof(str_length)
```
```{r}
accidents$Day <- (str_to_lower(accidents$Day))
```

```{r}
accidents$Vehicle.type <- str_to_upper(accidents$Vehicle.type)
```
```{r}
head(accidents)
```
```{r}
# function to turn all of the items in the dataset to lower 

lowerAll <- function (dataset) {
   for ( i in 1:ncol(dataset)) {
       if (class(dataset[,i])=="factor")
{dataset[,i] <- str_to_lower(dataset[,i])}
   }
   return(dataset)
}

```

```{r}
upperAll <- function(dataset) {
  for (i in 1: ncol(dataset)) {
    if (class (dataset[,i]) =="factor")
{dataset[,i] <- str_to_upper(dataset[,i])}
  }
  return(dataset)
}
```

```{r}
upperAll(accidents)
```


```{r}

non_alpha_num <- '[!\"#$%&\'()*+-,./:;<=>?@\\^_`{|}~\a\b\t\n\r\v\f]'

non_num <- paste(c(letters, LETTERS, non_alpha_num), collapse="")

num <- paste(c(0:9), collapse="")



```
```{r}
Num.post.caption 
```


```{r}
lowerAll(accidents)
```

```{r}
head(accidents,5)
```
```{r}
restrict_alpha_num <- gsub(".","", non_alpha_num, fixed=T)
restrict_alpha_num <- gsub("-","", restrict_alpha_num, fixed=T)

```



```{r}
#detecting bad characters 

bad_chars <-  str_detect(accidents$Response.difference, 
restrict_alpha_num)
any(bad_chars, na.rm=T)
sum(bad_chars)

```
```{r}
bad_factors <- str_subset(accidents$Response.difference, 
restrict_alpha_num)
bad_factors

```
```{r}
#2.	Obtain the list of Num.post.caption values which contain letters
#A, B or X. Tip: include your letters between '[ and ]'
```

```{r}
accidents$Response.difference <-str_remove_all(accidents$Response.difference,restrict_alpha_num)

```
```{r}
str_length("abcdef-g")
```

```{r}
grepl("f-", "abcdef-g")
```

```{r}
p1 <- str_locate("abcdef-g", "-" )
p1
```


```{r}
p2 <- str_locate("abcdef-g", "$" )
p2
```

```{r}
p3 <- str_locate("abcdef-g", "  " )
p3

```

```{r}
substring("abcdef-g", 3,6)
```

```{r}
levels(accidents$Age) <- c(levels(accidents$Age),"50-60","45-55", "30-40")

```
```{r}
accidents$Age[1] <- "50-60"
accidents$Age[5] <- "45-55"
accidents$Age[10] <- "30-40"

```





```{r}
findMean <- function (svalue){
   if(grepl("-", svalue)){ # find the position of the minus (if any)
      minus <- str_locate(svalue,"-")[1,1]
      num1 = as.numeric(substr(svalue, 0,minus-1)) # get 1st number
      num2 = as.numeric(substr(svalue,minus+1, nchar(svalue)))
      return(mean(c(num1,num2)))
   } else {return(as.numeric(svalue))} # no range specified
}

```


```{r}
accidents$Age <- sapply(as.character(accidents$Age),findMean)
```






```{r}
ageYoung <- function (svalue){
   if(grepl("y", svalue)){
      y <- str_locate(svalue,"y")[1,1] 
      num1 = as.numeric(substr(svalue, 0,y-1))
      m <- str_locate(svalue,"m") 
      num2 = as.numeric(substr(svalue,y+1,m-1))
      return(num1 + num2/12)
   } else {return(as.numeric(svalue))}
}

```


```{r}
ageYoung("3y4m")
```

```{r}
levels(accidents$Age) <- c(levels(accidents$Age),"2y3m","1y10m", "0y8m")
accidents$Age[100] <- "2y3m"
accidents$Age[570] <- "1y10m"

```

```{r}
accidents$Age[100] <- "2y3m"
accidents$Age[570] <- "1y10m"
accidents$Age[2000] <- "0y8m"

```

```{r}
accidents$Age <- sapply(as.character(accidents$Age),ageYoung)

accidents$Age
```
```{r}

```
3.	All report numbers start with 2015. Replace the report number values with the values obtained by removing 2015 from the beginning of the number.



4.	Some Num.postal.caption values contain a blank space in the middle. Replace the blank space by a minus sign.





5.	Column Cause…pedestrian contains a number of different values. Find out how many instances contain "pedestrian" in this column.





6.	Obtain all the records where the Num.postal.caption contains zeros in positions 6,7 and 8.



7.	Write a function which works with ages of the type "18m", i.e. where there is no year included, only months. Ensure you test your function.



8.	Introduce the following values in the Police.response.time attribute in any row you want: "<20", "<13", "~10",  "~18", "10mins", "12 mins",  "10-12", "10-15". Write code to clean this attribute. 



#Merging the new data 
```{r}
year <- c(2040,2045)
Oil <- c(2599.1, 2489.2)
Gas <- c(201.9, 193.4)
Coal <- c(24.7, 20.6)
Renewables <- c(125.8, 132.6)
newdata <- data.frame(year, Oil, Gas, Coal, Renewables,stringsAsFactors = T)

```

```{r}
colnames(consumption)
```

```{r}
#renaming a variable in r dplyr way 

consumption <- rename(consumption,  year = Million.tonnes.oil.equivalent)
```



```{r}
head(consumption)
```



```{r}
newConsumption <- rbind(consumption, newdata)
```


```{r}
#Retain only the rows where there are matching records in both datasets, i.e. the year appears in both datasets.
newConsumption2 <- merge(x = newConsumption, y = newdata,by.x = c("year"),by.y = c("year"),all.x = FALSE, all.y=FALSE)


```



```{r}
#Keep the records in newConsumption which don’t have a match in windAndSun. To do this, we set all.x to TRUE
newConsumption2 <- merge(x = newConsumption, y = newdata,by.x = c("year"),by.y=c("year"), all.x = TRUE, all.y=FALSE)

```




