---
title: "Lab 03"
author: "Chathura Perera"
date: "10/02/2022"
output: html_document
---

```{r}
accidents <- read.csv("https://www.dropbox.com/s/t68um5t7ajjdwjl/2015_accidents.csv?dl=1",header=T, stringsAsFactors = T)

```
```{r}




#the way i did it 
accidents_copy<- read.csv("https://www.dropbox.com/s/t68um5t7ajjdwjl/2015_accidents.csv?dl=1",header=T, stringsAsFactors = T)
```

```{r}
head(accidents)
```

```{r}
head(accidents)
```



```{r}
library(dplyr)
library(caret)
library(Hmisc)
library(ggplot2)
```


```{r}
#Removal of uneccassry 
accidents$Year <- NULL
accidents$Type.of.day <- NULL
accidents$Neighbourhood.name<-NULL
accidents$District.name<-NULL
accidents$Street.name<-NULL
head(accidents)
```

```{r}
accidents$District.code <- as.factor(accidents$District.code)
accidents$Neighbourhood.code <- as.factor(accidents$Neighbourhood.code)
accidents$Streeet.code <- as.factor(accidents$Streeet.code)
head(accidents,3)
```

```{r}
str(accidents)
```

```{r}
colnames(accidents)
```


```{r}
accidents$Age[accidents$Age == "Unknown"] <- NA
accidents$Age <- as.integer(as.character(accidents$Age))
```


```{r}
accidents$Response.difference[accidents$Response.difference == " "] <- NA
accidents$Response.difference[accidents$Response.difference == ">10"] <- NA
accidents$Response.difference[accidents$Response.difference == "~0"] <- NA
accidents$Response.difference[accidents$Response.difference == "nothing"] <- NA
summary(accidents$Response.difference)
```
```{r}
boxplot(accidents$Response.difference, col = "white")
stripchart(accidents$Response.difference,method = "overplot", jitter = 0.1,pch = 19,col = 4,vertical = TRUE,add = TRUE)            
#idea was to plot the points over the boxplot if you get it working let me know            
           
    
```
```{r}
#injury description name correction 
accidents$Injury.description[accidents$Injury.description == "minor injuery"] <-"minor injury" 
```
```{r}
summary(accidents$Injury.description)
```


```{r}
#Dropping levels and adding new levels
levels(accidents$Injury.description) 
```
```{r}
levels(accidents$Injury.description) <-droplevels(accidents$Injury.description)

```

```{r}
#issue check here 
levels(accidents$Injury.description) <-c(levels(accidents$Injury.description), "no injury")

```


```{r}
levels(accidents$in)
```
```{r}
cs <- preProcess(accidents, method = c("center","scale"))
standardisedAcc <- predict(cs, accidents)

```
```{r}
str(standardisedAcc)

```

```{r}
str(accidents)
```
```{r}
# proper way to make a copy 
accident <- accidents_copy
```



```{r}
ra <- preProcess(accidents, method = c("range"))
normalisedAcc <- predict(ra, accidents)

```

```{r}
str(normalisedAcc)
```



```{r}
dummies <- dummyVars( ~ ., accidents)
dummiesAccidents <- predict(dummies, accidents)

```

```{r}
str(dummiesAccidents)
typeof(dummiesAccidents)

```

In order to experiment with missing values, 
we need to introduce some missing values in our dataset. 
We introduce 20% missing values in column Response.difference 
and 15% in column Injury.description.


```{r}
accidents[sample(1:nrow(accidents), 0.2*nrow(accidents)), "Response.difference"] <- NA
accidents[sample(1:nrow(accidents), 0.15*nrow(accidents)), "Injury.description"] <- NA

```


```{r}
accidentsWithNa <- accidents
```

```{r}
apply(accidents, 2, function(x){return(sum(is.na(x)))})
```

```{r}
apply(accidentsWithNa, 2, function(x){return(sum(is.na(x)))})
```
```{r}
accidents <- accidents[complete.cases(accidents),]
```

```{r}
accidents$Response.difference <- impute((accidents$Response.difference), mean)

```
```{r}
accidents$Response.difference <- 
impute((accidents$Response.difference), median)

```

```{r}
accidents$Response.difference <- 
impute((accidents$Response.difference), "random")

```

FOR CATEGORICAL VALUES REPLACE WITH MODE 


```{r}
accidents$Injury.description <- 
impute((accidents$Injury.description), mode)

```

If NA is meaningful, you may wish to add it to the levels 
of the categorical column with missing values. For example, 
to add Naas a value for  the injury description column

```{r}
accidents$Injury.description <- addNA(accidents$Injury.description)
```


```{r}
knnImp <- preProcess(accidents, method = c("knnImpute"))
knnImputedAcc <- predict(knnImp, accidents)

```


```{r}
bagImp <- preProcess(accidents, method = c("bagImpute"))
bagimputedAcc <- predict(bagImp, accidents)

```

```{r}
head(bagimputedAcc)
```

```{r}
#filtering data 
quickAmbulance <-filter(accidents,  Ambulance.response.time <10)

```

```{r}
summary(quickAmbulance)
```

```{r}
quickAmbulanceDeath <-filter(accidents, Ambulance.response.time <10, 
Injury.description == "death")

```

```{r}
#APPROXIIMATIONS WITH NEAR 
around5minutes <- filter(accidents, 
near(5.0, Ambulance.response.time, tol=1.2))

```



```{r}
serious <-filter(accidents, (Injury.description == "serious injury" | Injury.description == "death")) &  (Ambulance.response.time > 30) )

```


